---
title: "Data Wrangling & Visualisation"
subtitle: 'Kelas Data R organised by <a href="https://www.instagram.com/mof.dac"> <i class="fab fa-instagram fa-fw"></i>mof.dac</a>'
author: 'Arif P. Sulistiono / @arifpras'
date: '28 October 2021 and 4 November 2021'
output:
  html_document:
    df_print: paged
    #self_contained: no
    toc: true
    toc_float: true
    theme: lumen #united
    highlight: tango
    fig_caption: yes
    css: style.css
    number_sections: TRUE
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

# Prerequisites

## Materials
Please visit https://github.com/arifpras/KelasData for the datasets and other relevant materials to be used during the sessions.

## Clearing the environment
```{r clearing-first, echo=TRUE}
rm(list=ls())
ls()
```

## Packages and libraries
```{r packages, include = FALSE}
#installing packages - all
#install.packages("tidyverse")
library(tidyverse)

#installing packages - separately
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("tidyr")
#install.packages("readxl")

#including when we call for library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)

#needs to be called individually
library(readxl)

#--install.packages("easypackages")
easypackages::libraries("dplyr", "ggplot2", "tidyr", "readxl")

#--install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, readxl)

#pacman::p_load(tidylog, warn.conflicts = FALSE)
#library(tidylog, warn.conflicts = FALSE)
#library(dplyr, warn.conflicts = FALSE)

#interactive plot
library(plotly)
```

```{r icon, include=FALSE}
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

```{r directory, echo=TRUE}
#which directory at the moment
getwd()

#setting working directory
setwd("/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData")

#dir()
```

## Datasets
```{r onepiece}
#sources:
##--https://www.kaggle.com/michau96
##--https://www.kaggle.com/aditya2803
##--https://www.opfanpage.com/2018/06/29/one-piece-power-ranking-chart/
##--https://listfist.com/list-of-one-piece-characters-by-age

#".csv" files - local

#op_chapters <- read.csv(file.choose())
#op_chapters <- read.csv("/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_chapters.csv")

#op_characters <- read.csv(file.choose())
#op_characters <- read.csv("/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_characters.csv")

#op_ratings <- read.csv(file.choose())
#op_ratings <- read.csv("/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_ratings.csv")

#op_powers <- read.csv(file.choose())
#op_powers <- read.csv("/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_powers.csv")

#op_ages <- read.csv(file.choose())
#op_ages <- read.csv("/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_ages.csv")

#".xlsx" files - local
#op_chapters <- read_excel(path = "/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_all.xlsx", sheet = "OP_chapters")
#op_characters <- read_excel(path = "/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_all.xlsx", sheet = "OP_characters")
#op_ratings <- read_excel(path = "/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_all.xlsx", sheet = "OP_ratings")
op_powers <- readxl::read_excel(path = "/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_all.xlsx", sheet = 4)
op_ages <- read_excel(path = "/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_all.xlsx", sheet = "OP_ages")

#".csv" files - cloud
#op_chapters <- read.csv("https://www.dropbox.com/s/t1vrxsa4ad0m2c4/OP_chapters.csv?dl=1")
#op_characters <- read.csv("https://www.dropbox.com/s/e74njw4iw7qfzbl/OP_characters.csv?dl=1")
#op_ratings <- read.csv("https://www.dropbox.com/s/qvdhlnkyk02giuj/OP_ratings.csv?dl=1")
#op_powers <- read.csv("https://www.dropbox.com/s/t3gbzzscz438ecp/OP_powers.csv?dl=1")
#op_ages <- read.csv("https://www.dropbox.com/s/r24vzv7eyz35irr/OP_ages.csv?dl=1")

#require(devtools)
#install_github("Displayr/flipAPI")
#library(flipAPI)

#".xlsx" files - cloud
#op_chapters <- flipAPI::DownloadXLSX("https://www.dropbox.com/s/5mev3xo0cd3tc61/OP_all.xlsx?dl=0", sheet = 1)
#op_characters <- flipAPI::DownloadXLSX("https://www.dropbox.com/s/5mev3xo0cd3tc61/OP_all.xlsx?dl=0", sheet = "OP_characters")
#op_ratings <- flipAPI::DownloadXLSX("https://www.dropbox.com/s/5mev3xo0cd3tc61/OP_all.xlsx?dl=0", sheet = 3)
#op_powers <- flipAPI::DownloadXLSX("https://www.dropbox.com/s/5mev3xo0cd3tc61/OP_all.xlsx?dl=0", sheet = "OP_powers", want.data.frame = TRUE)
#op_ages <- flipAPI::DownloadXLSX("https://www.dropbox.com/s/5mev3xo0cd3tc61/OP_all.xlsx?dl=0", sheet = "OP_ages", want.data.frame = TRUE)
```

```{r checking}
str(op_powers)
str(op_ages)

#or

glimpse(op_powers)
glimpse(op_ages)
```

***

# Data Wrangling

## Filtering
```{r filtering}
op_powers %>% 
  dplyr::filter(affiliations == "Marine")
```

## Arranging
```{r arranging}
op_powers %>%
  arrange(character) #ascending, alphabetical order
  #arrange(desc(character)) #descending, alphabetical order
```

## Recoding
```{r recoding}
law <- op_powers %>% 
  mutate(character02 = 
           recode(character, 
                  "Trafalgar D. Water Law" = "Trafalgar D. Law"))

law %>% filter(affiliations == "Heart Pirates") %>% 
  select(character, character02)
```

## Slicing
```{r slicing}
op_powers %>% 
  slice(18L) #variable number 18

op_powers %>% 
  slice(n()) #the last variable

op_powers %>% 
  slice(50:n())
```

```{r slice-min}
op_powers %>% 
  slice_min(character, n = 5, with_ties = FALSE)

op_powers %>% 
  slice_min(character, prop = 0.5, with_ties = TRUE)
```

```{r slice-tail}
op_powers %>% 
  slice_tail(n = 5)

op_powers %>% 
  slice_tail(prop = 0.5)
```

```{r slice-max}
op_powers %>% 
  slice_max(affiliations, n = 5, with_ties = FALSE)

op_powers %>% 
  slice_max(affiliations, prop = 0.5, with_ties = TRUE)
```

```{r slice-head}
op_powers %>% 
  slice_head(n = 5)

op_powers %>% 
  slice_head(prop = 0.5)
```

## Selecting
```{r selecting}
db03 <- op_powers %>%
  select(character, affiliations)
db03

db04 <- op_powers %>% 
  select(-level)
db04
```

## Relocating
```{r relocating}
db05 <- op_powers %>% 
  relocate(character, power, affiliations)
db05
```

## Renaming
```{r renaming}
status <- op_powers %>% 
  rename("status" = "level")
status
```

## Counting
```{r counting}
op_powers %>% 
  filter(affiliations == "Marine") %>% 
  count() %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

## Mutating
```{r mutating}
db06 <- op_powers %>% 
  mutate(anime = "One Piece",
         power_2 = power/2)

db06
```

```{r transmuting}
db07 <- op_powers %>% 
  transmute(anime = "One Piece",
            power_2 = power/2)

db07
```

## Grouping
```{r group-by}
db08 <- op_powers %>% 
  arrange(affiliations, desc(power)) %>% 
  group_by(affiliations) %>% 
  mutate(rank = row_number()) %>% 
  ungroup() %>% #close the group_by()
  arrange(affiliations) %>% 
  select(character, affiliations, rank)

db08
```

## Summarising
```{r summarising}
db09 <- op_powers %>% 
  arrange(affiliations, desc(power)) %>% 
  group_by(affiliations) %>% 
  summarise(avg = mean(power, na.rm = TRUE),
            n = n()) %>%
  ungroup()

db09
```

```{r row-wise}
db10a <- op_powers %>%
  mutate(fase1 = power*0.25,
         fase2 = power*0.75)

db10a

db10b <- db10a %>% 
  rowwise(character) %>%
  summarise(total = sum(fase1, fase2))

db10b
```

## Dataset: Villain
```{r df-villain}
villain_aff <- data.frame(
  character = c("Marshall D. Teach", "Charlotte Linlin", "Kaido", "Sakazuki Akainu", "Borsalino Kizaru"),
  affiliation = c("Blackbeard Pirates", "Big Mom Pirates", "Beasts Pirates", "Marine", "Marine")
)

character_power <- data.frame(
  name = c("Marshall D. Teach", "Kaido", "Sakazuki Akainu", "Monkey D. Dragon", "Silvers Rayleigh"),
  power = c(100, 100, 95, 99, 93)
)
```

## Joining
```{r left-join}
#mutating joins
villain_aff %>% 
  left_join(character_power, by = c("character" = "name")) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r right-join}
#mutating joins
villain_aff %>% 
  right_join(character_power, by = c("character" = "name")) %>%
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r inner-join}
#mutating joins
villain_aff %>% 
  inner_join(character_power, by = c("character" = "name"))  %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left") 
```

```{r full-join}
#mutating joins
villain_aff %>% 
  full_join(character_power, by = c("character" = "name"))  %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r semi-join}
#filtering joins
villain_aff %>% 
  semi_join(character_power, by = c("character" = "name")) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r anti-join}
#filtering joins
villain_aff %>% 
  anti_join(character_power, by = c("character" = "name")) %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

## Conditional "ifelse"
```{r ifelse}
(
  db11 <- op_powers %>%
    mutate(villain = ifelse(affiliations == "Blackbeard Pirates", "Villain",
                            "NA"))
)
```

## Dataset: Luffy
```{r df-luffy}
luffy00 <- tribble(
  ~arc, ~power,
  "1-Punk Hazard", 75,
  "2-Dressrosa", NA,
  "3-Zou", 80,
  "4-Whole Cake Island", 82,
  "5-Levely", NA
)

kableExtra::kbl(luffy00) %>%
    kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")

luffy00 <- luffy00 %>% 
  add_row(arc = "6- Wano Country",
          power = 85)

kableExtra::kbl(luffy00) %>%
    kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")


is.na(luffy00)

#data frame
colSums(is.na(luffy00))

mean(luffy00$power)
mean(luffy00$power, na.rm = TRUE)

#vector
which(is.na(luffy00))
sum(is.na(luffy00)) 
```

## Treating "N/A"
```{r filling}
luffy01 <- luffy00 %>%
  fill(power, .direction = "up") %>% 
  mutate(method = "fill")

kableExtra::kbl(luffy01) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r na-mean}
luffy02 <- luffy00 %>%
  mutate(
    power = replace(power, is.na(power), mean(power, na.rm = TRUE)),
    method = "mean"
    )

#luffy02$power[is.na(luffy02$power)] <- mean(luffy02$power, na.rm = TRUE)
kableExtra::kbl(luffy02) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r interpolating}
luffy03 <- luffy00 %>% 
  mutate(power = zoo::na.approx(power),
         method = "interpolate")

kableExtra::kbl(luffy03) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r dropping}
luffy00 %>% drop_na() %>% 
  kableExtra::kbl() %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

## Binding
```{r row-bind}
luffy04 <- rbind(luffy01, luffy02, luffy03)

luffy04 <- luffy01 %>%
  bind_rows(luffy02, luffy03)

kableExtra::kbl(luffy04) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r col-bind}
luffy05 <- cbind(luffy01, luffy02, luffy03)

luffy05 <- luffy01 %>% 
  bind_cols(luffy02, luffy03)

kableExtra::kbl(luffy05) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

## Long-table
```{r piv-long}
# selecting variables
(
  db01a <-  op_powers %>%
    left_join(op_ages, by = c("character" = "name")) %>%
    select(-level) %>%
    relocate(character, affiliations, power, age)
)

# pivoting
(
  db01b <- db01a %>%
    pivot_longer(cols = power:age,
                 names_to = "variable",
                 values_to = "value")
)
```

## Wide-table
```{r piv-wide}
(
  db01c <- db01b %>%
    pivot_wider(names_from = variable,
                values_from = value)
)
```

## Exporting Table
```{r exporting-table}
writexl::write_xlsx(x = list(db01aSheet = db01a, db01bSheet = db01b, db01cSheet = db01c), path = "OnePiece.xlsx")

write.csv(db01a, file = "csv_db01a.csv")
```

## Converting Characters
```{r mutate-each-upper}
luffy06 <- luffy04 %>% 
    mutate_each(method, funs = toupper)

kableExtra::kbl(luffy06) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r mutate-each-lower}
luffy07 <- luffy06 %>% 
    mutate_each(method, funs = tolower)

kableExtra::kbl(luffy07) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r mutate-if-upper}
luffy08 <- luffy07 %>%
  mutate_if(is.character, str_to_upper)

kableExtra::kbl(luffy08) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r mutate-if-lower}
luffy09 <- luffy08 %>%
  mutate_if(is.character, str_to_lower)

kableExtra::kbl(luffy09) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r mutate-all-upper}
luffy10 <- luffy09 %>% 
    mutate_all(., .funs = toupper)

kableExtra::kbl(luffy10) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
#drawback: converting numeric to character
```

```{r mutate-across-upper}
luffy11 <- luffy09 %>%
  mutate(across(where(is.character), 
                  toupper))

kableExtra::kbl(luffy11) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r mutate-across-lower}
luffy12 <- luffy11 %>% 
    mutate(across(where(is.character), 
                  tolower))

kableExtra::kbl(luffy12) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

```{r mutate-across-combine}
luffy13 <- luffy09 %>% 
    mutate(across(arc, tolower),
           across(method, toupper),
           across(power, round)
           )
kableExtra::kbl(luffy13) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

## Substituting
```{r eliminating-numeric}
luffy14 <- luffy13 %>% 
    mutate(
      power2 = gsub("[0-5]*", "", power), #taking out number 1 to 5
      arc2 = gsub(".*^", "0", arc), #adding 0 before the first number
      method2 = gsub("LL", "zz", method) #replacing LL to zz
      )

kableExtra::kbl(luffy14) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", 
                full_width = F, 
                position = "left")
```

***

# Data Visualisation

## Datasets

```{r datasets-ratings}
op_ratings <- read_excel(path = "/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_all.xlsx", sheet = "OP_ratings")

glimpse(op_ratings)
```

```{r filtering-ratings}
(
op_ratings02 <- op_ratings %>%
  select(episode, start, total_votes, average_rating) %>% 
  pivot_longer(cols = total_votes:average_rating,
               names_to = "variable",
               values_to = "value")
)
```

```{r shapes, eval=FALSE, include=FALSE}
#Source: https://github.com/tidyverse/ggplot2/issues/2075

pch_table <- c("0" = "square open",
               "1" = "circle open",
               "2" = "triangle open",
               "3" = "plus",
               "4" = "cross",
               "5" = "diamond open",
               "6" = "triangle down open",
               "7" = "square cross",
               "8" = "asterisk",
               "9" = "diamond plus",
               "10" = "circle plus",
               "11" = "star",
               "12" = "square plus",
               "13" = "circle cross",
               "14" = "square triangle",
               "15" = "square",
               "16" = "circle small",
               "17" = "triangle",
               "18" = "diamond",
               "19" = "circle",
               "20" = "bullet",
               "21" = "circle filled",
               "22" = "square filled",
               "23" = "diamond filled",
               "24" = "triangle filled",
               "25" = "triangle down filled")

df_shapes <- data.frame(shape = 0:25, shape_name = factor(paste0(0:25, " ('", pch_table, "')")))
df_shapes <- df_shapes[order(pch_table),]

ggplot(df_shapes, aes(0, 0, shape = shape)) +
  geom_point(aes(shape = shape), size = 5, fill = 'red', stroke = 2) +
  scale_shape_identity() +
  facet_wrap(~reorder(shape_name, shape)) +
  theme_void()

sort(pch_table)
```

## Palette: RColorBrewer
```{r rcolorbrewer}
pacman::p_load(RColorBrewer)

#RColorBrewer
display.brewer.all(type = "qual")
display.brewer.all(type = "seq")
display.brewer.all(type = "div")
#display.brewer.pal(n = 12, name = "Set2")
#brewer.pal(n = 12, name = "Set3")
#display.brewer.pal(n = 11, name = "RdBu")
```

## Palette: nord
```{r nord}
pacman::p_load(nord)

nord_show_palette("snowstorm")
nord_show_palette("frost")
nord_show_palette("aurora")
nord_show_palette("lumina")
nord_show_palette("mountain_forms")
nord_show_palette("silver_mine")
nord_show_palette("lake_superior")
nord_show_palette("victory_bonds")
nord_show_palette("halifax_harbor")
nord_show_palette("moose_pond")
nord_show_palette("algoma_forest")
nord_show_palette("rocky_mountain")
nord_show_palette("red_mountain")
nord_show_palette("baie_mouton")
nord_show_palette("afternoon_prarie")
#hex-code: https://www.nordtheme.com/docs/colors-and-palettes
```

## Palette: wesanderson
```{r wesanderson}
pacman::p_load(wesanderson)

#names(wes_palettes)
wes_palette("BottleRocket1")
wes_palette("BottleRocket2")
wes_palette("Rushmore1")
wes_palette("Royal1")
wes_palette("Royal2")
wes_palette("Zissou1")
wes_palette("Darjeeling1")
wes_palette("Darjeeling2")
wes_palette("Chevalier1")
wes_palette("FantasticFox1")
wes_palette("Moonrise1")
wes_palette("Moonrise2")
wes_palette("Moonrise3")
wes_palette("Cavalcanti1")
wes_palette("GrandBudapest1")
wes_palette("GrandBudapest2")
wes_palette("IsleofDogs1")
wes_palette("IsleofDogs2")

#hex-code: https://github.com/karthik/wesanderson/blob/master/R/colors.R
#scales::show_col(wes_palette("IsleofDogs2"))
```

## Palette: viridis
```{r viridis}
pacman::p_load(viridis)

scales::show_col(viridis_pal(option = "viridis")(6))
scales::show_col(viridis_pal(option = "plasma")(6))
scales::show_col(viridis_pal(option = "rocket")(6))
scales::show_col(viridis_pal(option = "magma")(6))
scales::show_col(viridis_pal(option = "rocket")(6))
scales::show_col(viridis_pal(option = "mako")(6))
scales::show_col(viridis_pal(option = "inferno")(6))
```

## Palette: Sequential
```{r sequential}
colorspace::hcl_palettes(type = "sequential", plot = TRUE) 
```

## Palette: Diverging
```{r diverging}
colorspace::hcl_palettes(type = "diverging", plot = TRUE, n = 9)
#colorspace::divergingx_palettes(plot = TRUE, n = 9)
```

## Colorblind Friendly
```{r Okabe-Ito}
(OkabeIto <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2",
             "#D55E00", "#CC79A7", "#000000"))
#https://colorschemedesigner.com/csd-3.5/

pals::pal.bands(OkabeIto)

#Source: https://jfly.uni-koeln.de/color/#pallet
#Source: https://davidmathlogic.com/colorblind/#%23D81B60-%231E88E5-%23FFC107-%23004D40
```

## Line chart: Base 
```{r line-base, out.width = "100%", fig.align = "center", fig.retina = 3}
(
line01 <- op_ratings02 %>% 
  ggplot(aes(x = episode, y = value)) +
  geom_line()
)
```

## Line chart: Faceting
```{r line-faceting, out.width = "100%", fig.align = "center", fig.retina = 3}
(
line02 <- op_ratings02 %>% 
  ggplot(aes(x = episode, y = value)) +
    geom_line() +
    facet_wrap(~variable, scales = "free", nrow = 2)
)
```

## Line chart: Color and Theme
```{r color-theme, out.width = "100%", fig.align = "center", fig.retina = 3}
(
line02 <- op_ratings02 %>%
  #modify the dataset
  filter(start > 2015 & start < 2021) %>%
  mutate(variable = recode(variable,
                           average_rating = "Average Rating",
                           total_votes = "Total Votes"))  %>%
  #plot the dataset
  ggplot(aes(x = episode, y = value)) +
  facet_wrap(~variable, scales = "free", nrow = 2) +
  geom_line(color = "#0072B2") + #colorblind-friendly
  theme_test() #could be explored even further
)
```

## Line chart: Coloring Based on Variable
```{r coloring-by-variable, out.width = "100%", fig.align = "center", fig.retina = 3}
(
line02 <- op_ratings02 %>%
  #modify the dataset
  filter(start > 2015 & start < 2021) %>%
  mutate(variable = recode(variable,
                           average_rating = "Average Rating",
                           total_votes = "Total Votes"))  %>%
  #plot the dataset
  ggplot(aes(x = episode, y = value)) +
  facet_wrap(~variable, scales = "free", nrow = 2) +
  geom_line(aes(color = start)) +
  theme_test() #could be explored even further
)
```

## Line chart: Colorblind Palette
```{r colorblind-palette, out.width = "100%", fig.align = "center", fig.retina = 3}
(
line02 <- op_ratings02 %>%
  #modify the dataset
  filter(start > 2015 & start < 2021) %>%
  mutate(variable = recode(variable,
                           average_rating = "Average Rating",
                           total_votes = "Total Votes"))  %>%
  #plot the dataset
  ggplot(aes(x = episode, y = value)) +
  facet_wrap(~variable, scales = "free", nrow = 2) +
  geom_line(aes(color = factor(start))) + 
  ggthemes::scale_color_colorblind() +
  theme_test() #could be explored even further
)
```

## Line Chart: Coloring Manual
```{r color-manual, out.width = "100%", fig.align = "center", fig.retina = 3}
(
line02 <- op_ratings02 %>%
  #modify the dataset
  filter(start > 2015 & start < 2021) %>%
  mutate(variable = recode(variable,
                           average_rating = "Average Rating",
                           total_votes = "Total Votes"))  %>%
  #plot the dataset
  ggplot(aes(x = episode, y = value)) +
  facet_wrap(~variable, scales = "free", nrow = 2) +
  geom_line(aes(color = factor(variable))) + 
  scale_color_manual(values = c("blue", "red")) +
  theme_test() #could be explored even further
)
```

## Line chart: Modification
```{r modification, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  line03 <- op_ratings02 %>%
    filter(start > 2015 & start < 2021) %>%
    mutate(
      variable = recode(variable,
                        average_rating = "Average Rating",
                        total_votes = "Total Votes")
    )  %>%
    ggplot(aes(x = episode, y = value)) +
    geom_line(aes(color = factor(start))) +
    #scale_color_brewer(palette = "Set1") +
    nord::scale_color_nord("aurora", discrete = TRUE) +
    facet_wrap(
      ~ variable,
      scales = "free",
      nrow = 2,
      strip.position = "right"
    ) +
    theme_test() +
    theme(
      legend.position = "bottom",
      plot.title = element_text(hjust = 0, size = 16, face = "bold"),
      plot.subtitle = element_text(hjust = 0, size = 14),
      plot.caption = element_text(hjust = 0, size = 8),
      plot.title.position = "plot",
      plot.caption.position = "plot"
    ) +
    labs(
      title = "One Piece",
      subtitle = "Ratings and Votes",
      caption = "\nSource: https://www.kaggle.com/aditya2803",
      x = "Episode",
      y = ""
    ) +
    guides(
      color = guide_legend(title = "Year release:")
    )
)
```

## Line chart: Adding points
```{r adding-points, out.width = "100%", fig.align = "center", fig.retina = 3}
(
line03 <- op_ratings02 %>% 
  filter(start > 2015 & start < 2021) %>%
  mutate(variable = recode(variable, average_rating = "Average Rating",
                           total_votes = "Total Votes")
         ) %>% 
  ggplot(aes(x = episode, y = value)) +
  facet_wrap(vars(variable), nrow = 2, scales = "free",
             as.table = TRUE, strip.position = "right") +
  geom_line(aes(color = factor(start))) +
  scale_color_brewer(palette = "Set1") +
  theme_classic() +
  #ggthemes::theme_economist() +
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0, size = 16,
                              face = "bold"),
    plot.subtitle = element_text(hjust = 0, size = 13),
    plot.caption = element_text(hjust = 0, size = 8),
    plot.title.position = "plot",
    plot.caption.position =  "plot",) +
  labs(title = "One Piece",
       subtitle = "Ratings and Votes",
       y = "",
       x = "Episode",
       caption = "\nSource: https://www.kaggle.com/aditya2803") +
  geom_point(aes(color = factor(start), fill = factor(start)), 
             shape = 21, size = 1, alpha = 0.5) +
  scale_fill_brewer(palette = "Set1") +
  guides(color = guide_legend(title = "Year release:"),
         fill = guide_legend(title = "Year release:"))
)
```

## Point chart: Dataset
```{r}
(
db02 <- db01c %>% 
  arrange(desc(power)) %>% 
  slice_max(power, n = 15, with_ties = TRUE)
)
```

## Point chart: Base
```{r point-base, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  point01 <- ggplot(db02, mapping = aes(x = character, y = age))
)
```
## Point chart: Flipping
```{r point-flip, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  point01 + coord_flip()
)
```
## Point chart: Assigning the function
```{r geom-point, out.width = "100%", fig.align = "center", fig.retina = 3}
point01 + 
  coord_flip() +
  geom_point()
```
## Point chart: Defining the color
```{r point-color, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  point01 + 
  coord_flip() +
  geom_point(color = "blue",
             shape = 0,
             size = 5)
)
```

## Point chart: Changing the shape
```{r point-shape, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  point01 + 
  coord_flip() +
  geom_point(color = "blue",
             shape = 21,
             size = 3)
)
```
## Point chart: Size
```{r point-size, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  point02 <- db02 %>% 
    ggplot(aes(character, age, size = power)) +
    coord_flip() +
    geom_point()
)
```

## Point chart: Reorder
```{r plot-reorder, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  point03 <- db02 %>% 
    ggplot(aes(reorder(character, power), age, size = power)) +
    coord_flip() +
    geom_point()
)
```

## Static Plot: No Faceting
```{r point04, out.width = "100%", fig.align = "center", fig.retina = 3}
(point04 <- op_powers %>%
  left_join(op_ages, by = c("character" = "name")) %>%
  relocate(character, age, power) %>%
  #group_by(level) %>%
  slice_max(power, n = 15, with_ties = FALSE) %>%
  #ungroup() %>%
  ggplot(aes(x = reorder(character, power),
             y = age)) +
  geom_point(aes(color = factor(affiliations), size = power)) +
  #coord_cartesian(expand = TRUE, clip = "on") +
  coord_flip() +
  nord::scale_color_nord("aurora") +
  xlab("") +
  ylab("Age") +
  theme_test() +
  theme(
    axis.text.x = element_text(size = 8),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.title.x = element_text(size = 7),
    axis.line.y = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 8),
    plot.title.position = "plot",
    plot.caption.position = "plot",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.key.size = unit(0.5, "cm"),
    legend.spacing = unit(1, "cm"),
    plot.title = element_text(hjust = 0, size = 13, face = "bold"),
    plot.subtitle = element_text(hjust = 0, size = 12),
    plot.caption = element_text(hjust = 0, size = 12),
    # legend.title = element_blank(),
    legend.position = "right",
    strip.text.x = element_text(size = 8),
    panel.grid.major.y = element_line(colour = "grey97"),
    panel.ontop = FALSE,
    panel.spacing = unit(1, "lines")
  ) +
  guides(size = guide_legend("Power"), color = guide_legend("Affiliation")) +
  labs(title = "The top 15-most powerful characters",
       subtitle = "",
       caption = "") +
  scale_size_continuous(range = c(1, 5)) +
  scale_y_continuous(limits = c(10, 90),
                     breaks = c(20, 50, 80))
)
```

## Static Plot: With Faceting
```{r point05, out.width = "100%", fig.align = "center", fig.retina = 3}
(point05 <- op_powers %>%
  left_join(op_ages, by = c("character" = "name")) %>%
  relocate(character, age, power) %>%
  #group_by(level) %>%
  slice_max(power, n = 15, with_ties = FALSE) %>%
  #ungroup() %>%
  ggplot(aes(x = reorder(character, power),
             y = age)) +
  geom_point(aes(color = factor(affiliations), size = power)) +
  coord_cartesian(expand = TRUE, clip = "on") +
  coord_flip() +
  nord::scale_color_nord("aurora") +
  xlab("") +
  ylab("Age") +
  theme_test() +
  theme(
    axis.text.x = element_text(size = 8),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.title.x = element_text(size = 7),
    axis.line.y = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 8),
    plot.title.position = "plot",
    plot.caption.position = "plot",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.key.size = unit(0.5, "cm"),
    legend.spacing = unit(1, "cm"),
    plot.title = element_text(hjust = 0, size = 13, face = "bold"),
    plot.subtitle = element_text(hjust = 0, size = 12),
    plot.caption = element_text(hjust = 0, size = 12),
    # legend.title = element_blank(),
    legend.position = "right",
    strip.text.x = element_text(size = 8),
    panel.grid.major.y = element_line(colour = "grey97"),
    panel.ontop = FALSE,
    panel.spacing = unit(1, "lines")
  ) +
  guides(size = guide_legend("Power"), color = guide_legend("Affiliation")) +
  labs(title = "The top 15-most powerful characters",
       subtitle = "",
       caption = "") +
  scale_size_continuous(range = c(1, 5)) +
  scale_y_continuous(limits = c(10, 90),
                     breaks = c(20, 50, 80)) +
  facet_grid(. ~ level)
)
```

## Saving Image
```{r saving-image}
ggsave(
  filename = "point05.png",
  #filename = "point05.pdf",
  plot = point05,
  #device = cairo_pdf,
  width = 297,
  height = 210,
  units = "mm"
)
```

## Interactive Plot: No Faceting
```{r point06, out.width = "100%", fig.align = "center", fig.retina = 3, fig.cap = 'Figure 3: The top 15-most powerful characters, without ***facet_grid(. ~ level)***', fig.topcaption = TRUE}
point06 <- op_powers %>%
  left_join(op_ages, by = c("character" = "name")) %>%
  relocate(character, age, power) %>%
  #group_by(level) %>%
  slice_max(power, n = 15, with_ties = FALSE) %>%
  #ungroup() %>%
  mutate(text = paste(
    "Character: ", reorder(character, power), 
    "\nAffiliation: ", factor(affiliations), 
    "\nAge: ", age, 
    "\nPower: ", power, 
    sep = "")) %>%
  ggplot(aes(x = reorder(character, power),
             y = factor(affiliations),
             text = text)) +
  geom_point(aes(color = age, size = power)) +
  coord_cartesian(expand = TRUE, clip = "on") +
  coord_flip() +
  #nord::scale_color_nord("victory_bonds", discrete = FALSE) +
  viridis::scale_colour_viridis(option = "viridis", direction = -1) +
  xlab("") +
  ylab("") +
  theme_test() +
  theme(
    axis.text.x = element_text(
      size = 8,
      angle = 90,
      vjust = 0.5,
      hjust = 1
    ),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.title.x = element_text(size = 7),
    axis.line.y = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 8),
    plot.title.position = "plot",
    plot.caption.position = "plot",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.key.size = unit(0.5, "cm"),
    legend.spacing = unit(1, "cm"),
    plot.title = element_text(hjust = 0, size = 13, face = "bold"),
    plot.subtitle = element_text(hjust = 0, size = 12),
    plot.caption = element_text(hjust = 0, size = 12),
    # legend.title = element_blank(),
    legend.position = "none",
    strip.text.x = element_text(size = 8),
    panel.grid.major.y = element_line(colour = "grey97"),
    panel.ontop = FALSE
  ) +
  guides(size = guide_legend("Power"), color = guide_colorbar("Age")) +
  labs(title = "",
       subtitle = "",
       caption = "") +
  scale_size_continuous(range = c(1, 5))
  
ggplotly(point06, tooltip = "text")
```

## Interactive Plot: With Faceting
```{r point07, out.width = "100%", fig.align = "center", fig.retina = 3, fig.cap = "Figure 4: The top 15-most powerful characters, with ***facet_grid(. ~ level)***", fig.topcaption = TRUE}
point07 <- op_powers %>% 
  left_join(op_ages, by = c("character" = "name")) %>% 
  relocate(character, age, power) %>% 
  #group_by(level) %>% 
  slice_max(power, n = 15, with_ties = FALSE) %>%
  #ungroup() %>% 
  mutate(text = paste(
    "Character: ", reorder(character, power), 
    "\nAffiliation: ", factor(affiliations), 
    "\nAge: ", age, 
    "\nPower: ", power, 
    sep = "")) %>%
  ggplot(aes(x = reorder(character, power), 
             y = factor(affiliations),
             text = text)) +
  geom_point(aes(color = age, size = power)) +
  coord_cartesian(expand = TRUE, clip = "on") +
  coord_flip() +
  #nord::scale_color_nord("victory_bonds", discrete = FALSE) +
  viridis::scale_colour_viridis(option = "viridis", direction = -1) +
  xlab("") +
  ylab("") +
  theme_test() +
  theme(
    axis.text.x = element_text(size = 8, angle = 90, vjust = 0.5, hjust=1),
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.title.x = element_text(size = 7),
    axis.line.y = element_blank(),
    axis.title.y = element_text(size = 7),
    axis.text.y = element_text(size = 8),
    plot.title.position = "plot",
    plot.caption.position = "plot",
    legend.title = element_text(size = 7),
    legend.text = element_text(size = 7),
    legend.key.size = unit(0.5, "cm"),
    legend.spacing = unit(1, "cm"),
    plot.title = element_text(
      hjust = 0, size = 13, face = "bold"
    ),
    plot.subtitle = element_text(hjust = 0, size = 12),
    plot.caption = element_text(hjust = 0, size = 12),
    # legend.title = element_blank(),
    legend.position = "none",
    strip.text.x = element_text(size = 8),
    panel.grid.major.y = element_line(colour = "grey97"),
    panel.ontop = FALSE,
    panel.spacing = unit(1, "lines")
  ) +
  guides(size = guide_legend("Power"), color = guide_colorbar("Age")) +
  labs(
    title = "",
    subtitle = "",
    caption = ""
  ) +
  scale_size_continuous(range = c(1,5)) +
  facet_grid(.~level)

ggplotly(point07, tooltip = "text")
```

***

# Clearing

```{r clearing-last, echo=TRUE}
#clearing the environment
ls()
rm(list=ls())
```

***

# Acknowledgements

## References

*   Grolemund, G., & Wickham, H. (2017). R for Data Science. O’Reilly Media.
*   R Core Team (2019). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
*   Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686.
*   Claus O. Wilke: Data Visualization in R, https://wilkelab.org/SDS375/
*   Cedric Scherer's personal blog: https://www.cedricscherer.com/
*   The R Graph Gallery: https://www.r-graph-gallery.com/index.html
*   ...and other sources i.e. stackoverflow.com, github.com

## Datasets
*   One Piece chapter and character: https://www.kaggle.com/michau96
*   One Piece anime rating: https://www.kaggle.com/aditya2803
*   One Piece power ranking: https://www.opfanpage.com/2018/06/29/one-piece-power-ranking-chart/
*   One Piece characters' age: https://listfist.com/list-of-one-piece-characters-by-age