---
title: "Basic Modelling"
subtitle: 'Kelas Data R organised by <a href="https://www.instagram.com/mof.dac"> <i class="fab fa-instagram fa-fw"></i>mof.dac</a>'
author: 'Arif P. Sulistiono / @arifpras'
date: '11 November 2021'
output:
  html_document:
    df_print: paged
    #self_contained: no
    toc: true
    toc_float: true
    theme: lumen #united
    highlight: tango
    fig_caption: yes
    css: style.css
    number_sections: TRUE
    includes:
      after_body: footer.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

# Prerequisites

## Materials
Please visit https://github.com/arifpras/KelasData for the datasets and other relevant materials to be used during the sessions.

## Clearing the environment
```{r clearing-first, echo=TRUE}
rm(list=ls())
ls()
```

## Packages and libraries
```{r packages, include = FALSE}
#install.packages("pacman")
pacman::p_load(tidyverse, readxl, skimr)
```

```{r icon, include=FALSE}
#could be ignored, related to uploading to website
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```

```{r directory, echo=TRUE}
#setting working directory
setwd("/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets")
```

## Datasets
```{r onepiece}
op_sales <- readxl::read_excel(path = "/Users/arifpras/OneDrive - The University of Nottingham/BB_KelasData/KelasData/00_Datasets/OP_all.xlsx", sheet = "OP_sales")
```

```{r checking}
glimpse(op_sales)

#change the date format: from <POSITX> to <date>
op_sales <- op_sales %>% 
  mutate(release_date = as.Date(release_date))

glimpse(op_sales)
```

```{r pivot-longer}
(
  db01 <- op_sales %>%
    pivot_longer(
      cols = pages:vixi_avg,
      names_to = "obsvar",
      values_to = "obsval"
    ) %>%
    mutate(
      obsvar = recode(
        obsvar,
        chapters = "Chapters",
        last_movie_days = "Days from Last Movie",
        pages = "Pages",
        sales_mio = "Total Sales (in millions)",
        vixi_avg = "VIX Index (in average)"
      )
    )
)
```

# Plotting

## Line chart
```{r line-chart, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  plot01 <- db01 %>%
    #plot the dataset
    ggplot(aes(x = volume, y = obsval)) +
    geom_line(aes(color = obsvar)) +
    facet_wrap( ~ obsvar, scales = "free", nrow = 2) +
    theme_light() +
    labs(x = "\nVolume", y = "") +
    #ggthemes::scale_color_colorblind() +
    scale_color_brewer(palette = "Set1") +
    theme(
      axis.text.x = element_text(size = 8),
      axis.ticks.x = element_blank(),
      axis.line.x = element_blank(),
      axis.title.x = element_text(size = 7),
      axis.line.y = element_blank(),
      axis.title.y = element_text(size = 7),
      axis.text.y = element_text(size = 8),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 7),
      legend.key.size = unit(0.5, "cm"),
      legend.spacing = unit(1, "cm"),
      plot.title = element_text(hjust = 0, size = 13, face = "bold"),
      plot.subtitle = element_text(hjust = 0, size = 12),
      plot.caption = element_text(hjust = 0, size = 12),
      # legend.title = element_blank(),
      legend.position = "bottom",
      strip.text.x = element_text(size = 8),
      panel.grid.major.y = element_line(colour = "grey97"),
      panel.ontop = FALSE,
      panel.spacing = unit(1, "lines")
    ) +
    guides(color = guide_legend(title = "Observable variables:"))
)
```

## Box plot
```{r box-plot, out.width = "100%", fig.align = "center", fig.retina = 3}
pacman::p_load(colorspace)
(
  plot02 <- db01 %>%
    #plot the dataset
    ggplot(aes(x = volume, y = obsval)) +
    geom_boxplot(aes(
      color = obsvar,
      fill = after_scale(desaturate(lighten(color, .7), .7))
      ),
      size = 1) +
    #geom_line(aes(color = obsvar)) +
    facet_wrap(~ obsvar, scales = "free", nrow = 2) +
    theme_light() +
    labs(
      title = "Boxplot",
      x = "\nVolume", y = "") +
    #ggthemes::scale_color_colorblind() +
    scale_color_brewer(palette = "Set1") +
    theme(
      axis.text.x = element_text(size = 8),
      axis.ticks.x = element_blank(),
      axis.line.x = element_blank(),
      axis.title.x = element_text(size = 7),
      axis.line.y = element_blank(),
      axis.title.y = element_text(size = 7),
      axis.text.y = element_text(size = 8),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 7),
      legend.key.size = unit(0.5, "cm"),
      legend.spacing = unit(1, "cm"),
      plot.title = element_text(hjust = 0, size = 16, face = "bold"),
      plot.subtitle = element_text(hjust = 0, size = 12),
      plot.caption = element_text(hjust = 0, size = 12),
      # legend.title = element_blank(),
      legend.position = "bottom",
      strip.text.x = element_text(size = 8),
      panel.grid.major.y = element_line(colour = "grey97"),
      panel.ontop = FALSE,
      panel.spacing = unit(1, "lines")
      ) +
    guides(color = guide_legend(title = "Observable variables:"))
)
```

## Histogram
```{r histogram, out.width = "100%", fig.align = "center", fig.retina = 3}
#pacman::p_load(colorspace)
(
  plot03 <- db01 %>%
    #plot the dataset
    ggplot(aes(x = obsval)) +
    geom_histogram(bins = 50, 
                   aes(
                     color = obsvar,
                     fill = after_scale(desaturate(lighten(color, .7), .7))
                     )) +
    facet_wrap(~ obsvar, scales = "free", nrow = 2) +
    theme_light() +
    labs(
      title = "Histogram",
      x = "\nRelease Date", y = "") +
    #ggthemes::scale_color_colorblind() +
    scale_color_brewer(palette = "Set1") +
    theme(
      axis.text.x = element_text(size = 8),
      axis.ticks.x = element_blank(),
      axis.line.x = element_blank(),
      axis.title.x = element_text(size = 7),
      axis.line.y = element_blank(),
      axis.title.y = element_text(size = 7),
      axis.text.y = element_text(size = 8),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      legend.title = element_text(size = 7),
      legend.text = element_text(size = 7),
      legend.key.size = unit(0.5, "cm"),
      legend.spacing = unit(1, "cm"),
      plot.title = element_text(hjust = 0, size = 16, face = "bold"),
      plot.subtitle = element_text(hjust = 0, size = 12),
      plot.caption = element_text(hjust = 0, size = 12),
      # legend.title = element_blank(),
      legend.position = "bottom",
      strip.text.x = element_text(size = 8),
      panel.grid.major.y = element_line(colour = "grey97"),
      panel.ontop = FALSE,
      panel.spacing = unit(1, "lines")
      ) +
    guides(color = guide_legend(title = "Observable variables:"))
)
```

# Descriptive statistics
## Summary: skim
```{r summ-skim}
op_sales %>% select(pages, chapters, sales_mio, last_movie_days, vixi_avg) %>% 
  skimr::skim()
```

## Summary: stargazer
```{r summ_stargazer}
pacman::p_load(pastecs, stargazer)

(
  summ01 <- op_sales %>%
    select(pages:vixi_avg) %>%
    rename("Chapters" = chapters,
           "Days from Last Movie" = last_movie_days,
           "Pages" = pages,
           "Total Sales (in millions)" = sales_mio,
           "VIX Index (in average)" = vixi_avg)
)

summ01 %>%
  as.data.frame() %>% 
  stargazer(type = 'text', out = "descsumm.txt", digits = 1) #flip = TRUE
```

## Summary: dplyr
```{r summ_dplyr}
(
  summ02 <- db01 %>%
    group_by(obsvar) %>%
    summarise(
      Obs = n(),
      Mean = mean(obsval, na.rm = TRUE),
      Std.Dev = sd(obsval, na.rm = TRUE),
      Min. = min(obsval, na.rm = TRUE),
      Max. = max(obsval, na.rm = TRUE),
      P25 = quantile(obsval, 0.25, na.rm = TRUE),
      P75 = quantile(obsval, 0.75, na.rm = TRUE),
      Var. = var(obsval, na.rm = TRUE)
  )
)
```

## Summary: parameters
```{r summ-param}
pacman::p_load(parameters)

(
  summ03 <- op_sales %>%
    select(pages:vixi_avg) %>%
    rename("Chapters" = chapters,
           "Days from Last Movie" = last_movie_days,
           "Pages" = pages,
           "Total Sales (in millions)" = sales_mio,
           "VIX Index (in average)" = vixi_avg) %>%
    describe_distribution()
)
```

# Correlations
```{r corr}
library(corrr)
(
  corr01 <- op_sales %>% 
    select(pages:vixi_avg) %>%
    rename("Chapters" = chapters,
           "Days from Last Movie" = last_movie_days,
           "Pages" = pages,
           "Total Sales (in millions)" = sales_mio,
           "VIX Index (in average)" = vixi_avg) %>% 
    correlate() %>% 
    shave() %>% 
    fashion()
)
```

# Unit Root Test
```{r unit-root}
pacman::p_load(tidyr, tseries)

(
  unrt01 <- db01 %>%
    group_by(obsvar) %>%
    summarise(
      box.pvalue = Box.test(obsval, lag=1, type="Ljung-Box")$p.value,
      box = Box.test(obsval, lag=1, type="Ljung-Box")$p.value<0.05,
      adf.pvalue = adf.test(obsval, alternative = "stationary")$p.value,
      adf = adf.test(obsval, alternative = "stationary")$p.value<0.05,
      kpss.pvalue=kpss.test(obsval)$p.value,
      kpss=kpss.test(obsval)$p.value>0.05
      )
)
```

# Simple Regressions
## Initial specifications
$$ \text{Sales}_t = \alpha_0 + \alpha_1 \text{Chapters}_t + \alpha_2 \text{LastMovie}_t + \alpha_3 \text{VIX index}_t + \upsilon_t  $$

$$ \text{Sales}_t = \beta_0 + \beta_1 \text{Pages}_t + \beta_2 \text{LastMovie}_t + \beta_3 \text{VIX index}_t + \epsilon_t  $$

## Modelling: base
```{r ols01-base}
ols01bs <- lm(sales_mio ~ chapters + last_movie_days + vixi_avg, 
                data = op_sales)

summary(ols01bs)
```

```{r ols02-base}
ols02bs <- lm(sales_mio ~ pages + last_movie_days + vixi_avg, 
                data = op_sales)

summary(ols02bs)
```

## Modelling: interaction
```{r ols03-interaction-ch}
ols03bs <- lm(sales_mio ~ chapters + last_movie_days + vixi_avg + 
                last_movie_days * vixi_avg, 
                data = op_sales)

summary(ols03bs)
```

```{r ols04-interaction-pg}
ols04bs <- lm(sales_mio ~ pages + last_movie_days + vixi_avg + 
                last_movie_days * vixi_avg, 
                data = op_sales)

summary(ols04bs)
```

## Modelling: lagging
```{r ols05-lag-ch}
ols05bs <- lm(sales_mio ~ chapters + lag(chapters, n = 1L) + 
                last_movie_days + vixi_avg + last_movie_days * vixi_avg, 
                data = op_sales)

summary(ols05bs)
```

```{r ols06-lag-ch}
ols06bs <- lm(sales_mio ~ pages + lag(pages, n = 1L) + 
                last_movie_days + vixi_avg + last_movie_days * vixi_avg, 
                data = op_sales)

summary(ols06bs)
```

## Modelling: 1st difference
```{r ols07-diff-ch}
ols07bs <- lm(diff(sales_mio, lag = 1) ~ diff(chapters, lag = 1) + 
                diff(last_movie_days, lag = 1) + diff(vixi_avg, lag = 1), 
                data = op_sales)

summary(ols07bs)
```

```{r ols08-diff-pg}
ols08bs <- lm(diff(sales_mio, lag = 1) ~ diff(pages, lag = 1) + 
                diff(last_movie_days, lag = 1) + diff(vixi_avg, lag = 1), 
                data = op_sales)

summary(ols08bs)
```

## Modelling: dplyr
```{r ols01-dplyr}
pacman::p_load(broom)
(
  ols01dp <- op_sales %>%
    select(pages:vixi_avg) %>%
    lm(sales_mio ~ chapters + last_movie_days + vixi_avg, 
       data = .) %>%
    tidy()
)
```

```{r ols02-dplyr}
(
  ols02dp <- op_sales %>%
    select(pages:vixi_avg) %>%
    lm(sales_mio ~ pages + last_movie_days + vixi_avg, 
       data = .) %>%
    tidy()
)
```

## Modelling: estimatr
```{r ols01-estimatr}
pacman::p_load(estimatr)

(
  ols01es <- op_sales %>%
    lm_robust(sales_mio ~ chapters + last_movie_days + vixi_avg,
              data = . ,se_type = "stata") %>%
    tidy()
)
```

```{r ols02-estimatr}
pacman::p_load(estimatr)

(
  ols02es <- op_sales %>%
    lm_robust(sales_mio ~ pages + last_movie_days + vixi_avg,
              data = . ,se_type = "stata") %>%
    tidy()
)
```

## Export: stargazer
```{r report-strgz}
stargazer(ols01bs, ols02bs, type = "text", out = "ols_base.txt")
```

## Export: texreg
```{r report-texreg, out.width = "100%", fig.align = "center", fig.retina = 3}
pacman::p_load(texreg)

screenreg(list(ols01bs, ols02bs), digits = 2)
#knitreg(list(ols01bs, ols02bs), digits = 2)
#wordreg(list(ols01bs, ols02bs), digits = 2, file = "ols_reg.doc")
plotreg(list(ols01bs, ols02bs), digits = 2)
```

## Export: `lm_robust` & `wordreg`
```{r estm-strgz}
#doesn't work: lm_robust & stargazer & texreg
ols03es <- lm_robust(sales_mio ~ chapters + last_movie_days + vixi_avg,
              data = op_sales, se_type = "stata")

ols04es <- lm_robust(sales_mio ~ pages + last_movie_days + vixi_avg,
              data = op_sales, se_type = "stata")

#stargazer(ols03es, ols04es, type = "text", out = "ols_estimatr.txt")
#wordreg(list(ols03es, ols04es), digits = 2, file = "ols_estimatr.doc")
```

## Plot: sjPlot single
```{r sjPlot-single, out.width = "100%", fig.align = "center", fig.retina = 3}
#source: http://www.strengejacke.de/sjPlot/index.html
pacman::p_load(sjPlot, sjmisc, sjlabelled)
theme_set(theme_test())

plot_model(ols01bs,
           title = "Sales (in mio)",
           show.values = TRUE, 
           type = "est", 
           digits = 4) +
  #edit with ggplot2
  scale_y_continuous(limits = c(-0.12,0.12))

plot_model(ols02bs,
           title = "Sales (in mio)",
           show.values = TRUE, 
           type = "est", 
           digits = 4) +
  #edit with ggplot2
  scale_y_continuous(limits = c(-0.12,0.12))
```

## Plot: sjPlot multiple
```{r sjPlot-multiple, out.width = "100%", fig.align = "center", fig.retina = 3}
plot_models(ols01bs, ols02bs,
            title = "Sales (in mio)",
            show.values = TRUE, 
            show.intercept = FALSE, 
            line.size = 0.1,
            p.shape = TRUE,
            value.size = 3,
            digits = 4) +
  #edit with ggplot2
  scale_y_continuous(limits = c(-0.12,0.12)) +
  theme(
      axis.text.x = element_text(size = 8),
      axis.ticks.x = element_blank(),
      axis.line.x = element_blank(),
      axis.title.x = element_text(size = 8),
      axis.line.y = element_blank(),
      axis.title.y = element_text(size = 8),
      axis.text.y = element_text(size = 8),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      legend.title = element_text(size = 8),
      legend.text = element_text(size = 8),
      legend.key.size = unit(0.5, "cm"),
      legend.spacing = unit(1, "cm"),
      plot.title = element_text(hjust = 0, size = 16, face = "bold"),
      plot.subtitle = element_text(hjust = 0, size = 12),
      plot.caption = element_text(hjust = 0, size = 12),
      # legend.title = element_blank(),
      legend.position = "bottom",
      strip.text.x = element_text(size = 8),
      panel.grid.major.y = element_line(colour = "grey97"),
      panel.ontop = FALSE,
      panel.spacing = unit(1, "lines")
      )
```

## Plot: `forestmodel`
```{r forest, out.width = "100%", fig.align = "center", fig.retina = 3}
library(forestmodel)

forest_model(ols01bs, theme = theme_forest())
forest_model(ols02bs, theme = theme_forest())
```

## Performance: glance
```{r glancing}
gl01bs <- broom::glance(ols01bs) %>% t() %>% round(digits = 2)
gl02bs <- broom::glance(ols02bs) %>% t() %>% round(digits = 2)

cbind(gl01bs, gl02bs)
```

## Performance: check model
```{r check-mod01, out.width = "100%", fig.align = "center", fig.retina = 3}
pacman::p_load(patchwork, performance)
(
  ch01 <- check_model(ols01bs)
)
```

```{r check-mod02, out.width = "100%", fig.align = "center", fig.retina = 3}
(
  ch02 <- check_model(ols02bs)
)
```

## Performance: compare
```{r performance, out.width = "100%", fig.align = "center", fig.retina = 3}
#install.packages("performance", repos = "https://easystats.r-universe.dev")
library(performance)
library(see)

(
  comp <- compare_performance(ols01bs, ols02bs)
)
plot(comp)
```

# Clearing
```{r clearing-last, echo=TRUE}
#clearing the environment
ls()
rm(list=ls())
```

***

# Acknowledgements

## References

*   Grolemund, G., & Wickham, H. (2017). R for Data Science. O’Reilly Media.
*   R Core Team (2019). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
*   Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686.
*   Claus O. Wilke: Data Visualization in R, https://wilkelab.org/SDS375/
*   Cedric Scherer's personal blog: https://www.cedricscherer.com/
*   Kim, A. Y. and Ismay, C. and Kuhn, M. (2021) Take a moderndive into introductory linear regression with R Journal of Open Source Education. 4(41), 115 https://doi.org/10.21105/jose.00115
*   The R Graph Gallery: https://www.r-graph-gallery.com/index.html
*   ...and other sources i.e. stackoverflow.com, github.com

## Datasets
*   One Piece by volume: https://listfist.com/list-of-one-piece-volumes
*   One Piece by chapter: https://listfist.com/list-of-one-piece-manga-chapters
*   One Piece sales: https://erzat.blog/top-sales-according-to-series-for-the-month-of-september-2021/ and  https://twitter.com/WSJ_manga/status/1214168838511702016
*   VIX index: https://finance.yahoo.com/quote/%5EVIX?p=%5EVIX